---
title: "Analysis of R Packages on Stack Overflow Over Time"
output: html_document
---

### Read the data and process it

Note that this will take a few minutes to run as the datasets are rather large.

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

```{r, echo=FALSE}
library(tidyverse)
library(stringr)
library(scales)
library(dplyr)
library(knitr)

# Load pre-computed dataframe
setwd("c:/Users/jflam/src/StackOverflowR/StackOverflowR")
r_packages <- readRDS(file = "r_packages.Rda")
package_by_year <- readRDS(file = "package_by_year.Rda")
```

Let's take a look at the first few rows of the `r_packages` dataframe:

```{r, results='asis'}
library(magrittr)
library(kableExtra)
r_packages %>%
    top_n(10) %>%
    kable(format = "html") %>%
    kable_styling()
```

Let's look at the most mentioned R Packages in Stack Overflow Q&A

```{r}
r_packages %>%
    count(Package, sort = TRUE) %>%
    head(20) %>%
    mutate(Package = reorder(Package, n)) %>%
    ggplot(aes(Package, n)) +
    geom_col(fill = "#f48024") +
    coord_flip() +
    scale_y_continuous(labels = comma_format()) +
    labs(x = "",
       y = "# of mentions in Stack Overflow questions and answers",
       title = "Most Mentioned R Packages in Stack Overflow Q&A",
       subtitle = "In non-deleted questions and answers up to September 2017.")
```

Now, let's take those same packages, and look at the Q&A traffic over time

```{r}
package_by_year %>%
    filter(PackageTotal >= 1200) %>%
    mutate(Percent = n / YearTotal) %>%
    complete(Package, Year, fill = list(Percent = 0)) %>%
    mutate(Package = reorder(Package, - PackageTotal, mean)) %>%
    ggplot(aes(Year, Percent, color = Package)) +
    geom_line(show.legend = FALSE, size = 1) +
    facet_wrap(~Package) +
    scale_x_continuous(breaks = seq(2009, 2017, 2)) +
    scale_y_continuous(labels = percent_format()) +
    theme_light() +
    labs(x = "Year",
       y = "% of R questions where either the question or an answer uses package",
       title = "Use of R packages in Stack Overflow Q&A over time",
       subtitle = "For the 20 most-mentioned packages. 2017 includes only Jan-Sep.")
```

Now let's look at the Ecosystem of R packages based on correlations between different tags:

```{r}
options(repr.plot.width = 13, repr.plot.height = 8)

library(dplyr)
library(widyr)

package_posts <- r_packages %>%
  filter(PostType == "Answer") %>%
  distinct(ParentId, Package) %>%
  add_count(Package) %>%
  filter(n >= 100)

cors_filtered <- package_posts %>%
  pairwise_cor(Package, ParentId, sort = TRUE) %>%
  filter(correlation > .05)

package_counts <- r_packages %>%
  count(Package) %>%
  filter(Package %in% cors_filtered$item1 | Package %in% cors_filtered$item2)

vertices <- package_counts %>%
  inner_join(package_growth, by = "Package")

library(ggraph)
suppressPackageStartupMessages(library(igraph))

set.seed(2017)

cors_filtered %>%
  graph_from_data_frame(vertices = vertices) %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation), show.legend = FALSE) +
  geom_node_point(aes(size = Y2017 * 1.05), color = "black") +
  geom_node_point(aes(size = Y2017, color = pmin(4, Growth))) +
  scale_size_continuous(range = c(.5, 14), labels = percent_format(),
                        breaks = seq(.05, .2, .05)) +
  geom_node_text(aes(label = name), repel = TRUE, size = 4) +
  theme_void(base_size = 16) +
  scale_color_gradient2(low = "#0077CC", high = "#f48024", trans = "log", breaks = 2 ^ seq(-1, 2),
                        labels = c(".5X", "Same", "2X", "4X")) +
  labs(title = "Ecosystem of R packages",
       subtitle = "Correlations are based on packages often used in Stack Overflow answers on the same question.",
       size = "% of 2017 questions",
       color = "2017/2016 Growth") +
  expand_limits(x = -6)
```
